---
title: "Lab 1_KWH Version"
author: "Kynan Witters Hicks"
date: "January 10, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r library}

library(tidyverse) #Collection of data wrangling and visualization
library(sf)
library(leaflet)
library(tmap)
```

2. Review: reading, exploring, wrangling, and visulaizing data

a. Read in data
```{r}
wine <- read_csv("wine_data.csv")
```

b.Wrangling
```{r}

wine_new <- wine %>%
  select(country, province, winery, region_1, points, price) %>% 
  rename(state = province) %>% #stuck
  filter(state == "California" | state =="Oregon" | state == "Washington" ) %>%  
  mutate(ppd = points/price) %>% 
  arrange(-ppd) 
  
  
  #31: if put a minus sign in any of columns it means exclude column
  #32:"=" sets equal and "==" searches for a match
  #33: "|" means or
  #35: sorts by column decending
```

c. Use group_by() and summarize() to find summary statistics by group
```{r}

wine_summary <- wine_new %>% 
  group_by(state) %>% 
  summarize(
    mean_ppd = mean(ppd, na.rm = TRUE), 
    max_ppd = max(ppd, na.rm = TRUE)
  )

#48: Recognize every unique level and have those levels operate as unique groups
#50: "mean_ppd" is column name and "na.rm" is remove NA
#You can also group by more than one variable - group_by(state, type)
```

d. A few gg-graphics
```{r}

#anytime doing stat analysis create a histogram 
#always tell ggplot what data is and what x and y variables are and what type of graph we want to create

wine_hists <- ggplot(wine_new, aes(x = ppd)) +
  geom_histogram(aes(fill = state)) +
  facet_wrap(~state, scales = "free")

wine_hists

#66: always need to have "aes" if want to create fill based on variable
#67: split up as a function of (~) state. "Scales = free" sets independent scales for each graph

```

new graph: scatterplot graph of price vs points, with color determined by state, adding a linear model trendline on top of it

```{r}

wine_points <- ggplot(wine_new, aes(x = price, y = points)) + 
  geom_point(aes(color = state), alpha = 0.5) +
  geom_smooth(color = "black", method = "lm", se = FALSE) +
  scale_color_manual(values = c("royalblue",  "mediumorchid1", "orange")) +
  facet_wrap(~state) +
  theme_light()

wine_points

#geom_point use "color"" while geom_histogram use "fill"" for color
#82: override default colors
```

2. California counties (spatial data) using sf

sf is catching up to spatial data analysis in other programs quickly 
```{r}

ca_counties <- st_read(dsn = ".", layer = "CA_Counties_TIGER2016")

#plot(ca_counties)

#98: st_read is for spacial/temporal. "." means "we're already there"
#sf uses dataset as normal data frame and the geometry just sticks to it! This is why it's so cool. 

ca_land <- ca_counties %>% 
  select(NAME, ALAND)

ca_pop_inc <- read.csv("ca_pop_inc.csv") %>% 
  rename(NAME = COUNTY)

#we want to join the ca_pop_inc file to our spatial ca_counties files
#109: reanames COUNTY coloumn to NAME so that we can match columns in two datasets 

ca_df <- full_join(ca_land, ca_pop_inc) %>% 
  select(NAME, MedFamilyIncome)

#114: full_join is safest bet over "join" function because won't get rid of anything


```

d. Use geom_sf to make a map in ggplot

```{r}

ca_income <- ggplot(ca_df) + 
  geom_sf(aes(fill = MedFamilyIncome))

ca_income

#Now we want to overlay on a different map using "leaflet" package but have to change coordinate system first

ca_df_transform <- st_transform(ca_df, crs = 4326)

#133: 4326 is the corrdinate reference system 

leaflet(ca_df_transform) %>% 
  addTiles() %>% 
  addPolygons() 

#leaflet creates interactive set of polygons in Viewer

```

